---
title: "Assignment3"
author: "Mary Geschwindt"
date: "9/23/2020"
output: 
  html_document:
    theme: paper
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First I will load my libraries.

```{r message=FALSE}
library(sf)
library(tidyverse)
library(ggthemes)
```

## Geospatial Data

### Loading the Data

I have selected polygons and points from Open Data NYC. My spatial data includes:

A. Neighborhood boundaries (polygon)
B. Parks (polygon)
C. Bicycle parking spots/stations (point)
D. Parking meters (point)

```{r message=FALSE}
nhoods <- st_read("https://data.cityofnewyork.us/api/geospatial/cpf4-rkhq?method=export&format=KML", quiet = TRUE)

parks <- st_read("https://data.cityofnewyork.us/api/geospatial/g84h-jbjm?method=export&format=KML", quiet = TRUE)

bike_parking <- st_read("https://data.cityofnewyork.us/api/geospatial/yh4a-g3fj?method=export&format=KML", quiet = TRUE)

parking <- st_read("https://data.cityofnewyork.us/api/geospatial/5jsj-cq4s?method=export&format=KML", quiet = TRUE)
```

### Transforming the Data

Using a transverse Mercator projection (from NAD27) from spatialreference.org, I will now transform my data to this projection.

```{r message=FALSE}
NY_state_tmerc <- "+proj=tmerc +lat_0=40 +lon_0=-74.33333333333333 +k=0.999966667 +x_0=152400.3048006096 +y_0=0 +ellps=clrk66 +datum=NAD27 +to_meter=0.3048006096012192 +no_defs"

nhoods <- nhoods %>%
  st_transform(NY_state_tmerc)

parks <- parks %>%
  st_transform(NY_state_tmerc)

bike_parking <- bike_parking %>%
  st_transform(NY_state_tmerc)

parking <- parking %>%
  st_transform(NY_state_tmerc)
```

I'll create a test map:

```{r message=FALSE}
ggplot(nhoods) +
  geom_sf(fill = NA, color = "gray") +
  geom_sf(data = )
```

