---
title: "Assignment 6 Interactive Map"
author: "Mary Geschwindt"
date: "10/21/2020"
output:
  html_document:
    theme: paper
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment Setup

### Loading Libraries

I'll use the following libraries for this assignment:

```{r load libraries, message=FALSE}
library(tidyverse)
library(sf)
library(leaflet)
library(htmltools)
library(htmlwidgets)
library(ggplot2)
library(ggthemes)
library(units)
```

### Loading Data: Bicycle Parking Point Data

I'm excited about this exercise, because there is a map I have always wished existed: a map showing all bike parking racks in New York City. I never considered that this would be a map that I could create myself, so I'm looking forward to attempting this.

I will bring in point data for the bicycle parking racks. My data source is the New York City Open Data portal. The dataset includes bicycle parking racks throughout New York City as of May 2019. The dataset alone has 26282 observations of existing bike racks; I will also include a filter to select only the borough of Manhattan for this exercise, which has only 8910 bicycle racks.

I will give credit to my classmate Megan Willis-Jackson who helped me to transform my data to the correct coordinate plane.

```{r load bike data, message=FALSE, results='hide'}
leaflet_plane <- "+proj=longlat +datum=WGS84"

bikeracks <- st_read("https://data.cityofnewyork.us/api/geospatial/yh4a-g3fj?method=export&format=GeoJSON") %>%
  filter(borough == "Manhattan") %>%
  st_transform(crs = leaflet_plane)

bikelanes <- st_read("https://data.cityofnewyork.us/api/geospatial/7vsa-caz7?method=export&format=GeoJSON") %>%
  filter(boro == "1") %>%
  st_transform(crs = leaflet_plane)
```

## Creating the Interactive Map

For my map, I would like to create points as circles to represent each bicycle parking rack in Manhattan. Each of these circles will also have a popup label which includes the name of the street on which the bike is located.

```{r create map, message=FALSE}
bikeracks_map <- leaflet(bikeracks,
                         options = leafletOptions(minZoom = 0.1, maxZoom = 0.8)) %>%
  addProviderTiles("Stamen.TonerLite") %>%
  addCircles(fillColor = "cyan", color = "cyan",
             stroke = FALSE, radius = 200,
             highlightOptions = highlightOptions(fillColor = "mediumblue",
                                                 fillOpacity = 1),
             popup = paste("Street Location:", bikeracks$street_nam)) %>%
  addPolylines(data = bikelanes, stroke = T,
               color = "pink",
               opacity = 1,
               weight = 1,
               highlightOptions = highlightOptions(color = "violet",
                                                   fillOpacity = 1)) %>%
  addControl("Bicycle Parking Racks in Manhattan", position = "topright") %>%
  addControl('<a href="https://data.cityofnewyork.us/api/geospatial/yh4a-g3fj?method=export&format=GeoJSON">Data source</a>',
             position = "bottomleft") %>%
  setMaxBounds(lng1 = 40.49804,
               lat1 = -74.25185,
               lng2 = 40.91142,
               lat2 = -73.70279)

bikeracks_map
```

### Saving HTML

```{r save widget, message=FALSE}
#saveWidget(bikeracks_map, file = "bikeracks_map.html")
```


## Assignment Conclusion

I may wish to further this develop this map to become something useful for the Bike NYC community. A further iteration would include listing all bicycle parking racks throughout the five boroughs. If such data is available, I imagine I could also create another layer which includes data of proposed bicycle rack locations; I would make this points a different color from the points showing the existing bike parking racks.

It's also important to include the data showing the existing (and potentially proposed) bicycle infrastructure. This would allow for several uses:

My intended audience is members of Bike NYC (anyone who is a part of the informal bicycling community). They would be able to use this map as part of their trip planning. It would inform bicyclists where there are bike parking racks located near their final destinations, and then plan routes directly to those bike racks.

Another means by which this map could be interpreted is by people making decisions about transportation infrastructure in New York City. Such a map would allow NYC DOT planners, City Council Members, and Community Boards to visually see where bicycle infrastructure is lacking, and highlight areas to study which may require more bike parking racks. For example, there may be a portion of the map which shows a protected bike parking lane on a route where there are many local businesses, yet there is sparse bike parking located along that route. Local leaders could view this map to determine that it may be in the neighborhood's best interest to include more bike parking, which allows for more opportunities that bicyclists will frequent those destinations.

A very, far future example of this map could be data that shows the bicycle racks in use in real-time. However, this would require some sort of real-time recording and updating of existing bicycle racks to include such technology where occupancy could be recorded.