---
title: "Assignment 7: Raster Data"
author: "Mary Geschwindt"
date: "10/29/2020"
output:
    html_document:
    theme: paper
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Document Setup

### Load Libraries

```{r load libraries, message=FALSE}
library(tidyverse)
library(sf)
library(leaflet)
library(htmltools)
library(htmlwidgets)
library(raster)
library(gstat)
library(spatial)
library(stringr)
```

### Load Data

I will load in the census tract level data for means of transportation to work and travel time to work for census tracts in Boston. I will join this to each census tract to create a dataset which reads the data for each tract and can apply it spatially. I will give credit to my classmate Aron Lessor who taught me how to do this dataset join for census level data. I am using the ACS from 2018.

```{r load data and create new dataset through join, message=FALSE, warning=FALSE}
tracts_info <- st_read("suffolk_county_acs2018_commuting.csv", quiet = TRUE) %>%
  dplyr::select(GEO_ID, S0802_C01_090E) %>%
  mutate(GEOID10 = as.numeric(str_sub(GEO_ID, -11, -1))) %>%
  dplyr::select(GEOID10, S0802_C01_090E)

census_tracts <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/4a8eb4fb3be44ed5a1eec28551b9f3b2_0.geojson?outSR=%7B%22latestWkid%22%3A3857%2C%22wkid%22%3A102100%7D", quiet = TRUE) %>%
  mutate(GEOID10 = as.numeric(GEOID10))

avg_travel_time <- left_join(census_tracts, tracts_info, by = "GEOID10", copy = FALSE) %>%
  mutate(S0802_C01_090E = case_when(S0802_C01_090E == "-" ~ NA_real_,
                                            S0802_C01_090E != "-" ~ as.numeric(S0802_C01_090E))) %>%
  mutate(S0802_C01_090E = case_when(is.na(S0802_C01_090E) ~ round(mean(S0802_C01_090E, na.rm = TRUE),1),
                                          !is.na(S0802_C01_090E) ~ S0802_C01_090E)) %>%
  mutate(S0802_C01_090E = case_when(S0802_C01_090E == "N" ~ NA_real_,
                                            S0802_C01_090E != "-" ~ as.numeric(S0802_C01_090E))) %>%
  mutate(S0802_C01_090E = case_when(is.na(S0802_C01_090E) ~ round(mean(S0802_C01_090E, na.rm = TRUE),1),
                                          !is.na(S0802_C01_090E) ~ S0802_C01_090E))

avg_travel_time = avg_travel_time[-48,]
```


## Chloropleth Map

```{r chloropleth map, message=FALSE}
avg_travel_time$describe <- paste("Tract", avg_travel_time$OBJECTID, "<br>", "Average Travel Time to Work", "Minutes", prettyNum(avg_travel_time$S0802_C01_090E, digits = 3)) %>%
  lapply(htmltools::HTML)

bins <- seq(min(avg_travel_time$S0802_C01_090E),
            max(avg_travel_time$S0802_C01_090E), by = 15)
pal <- colorNumeric("viridis",
                    domain = avg_travel_time$S0802_C01_090E,
                    na.color = "#00000000")

tt_chloropleth_map <- leaflet(avg_travel_time) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addPolygons(highlightOptions = highlightOptions(fillOpacity = 1),
              label = avg_travel_time$describe,
              fillColor = ~pal(S0802_C01_090E),
              weight = 1, color = "black") %>%
  addLegend(pal = pal,
            values = ~S0802_C01_090E,
            bins = 10,
            opacity = 0.7, title = "Average Travel Time to Work (Minutes)",
            position = "topright")

tt_chloropleth_map
```


## Points Map

```{r points map, message=FALSE, warning=FALSE}
MA_state_plane <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs "

WGS84 <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"

travel_time_points <- st_centroid(
  st_transform(avg_travel_time, crs = MA_state_plane)) %>%
  st_transform(WGS84)

tt_points_map <- leaflet(travel_time_points) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addCircles(highlightOptions = highlightOptions(fillOpacity = 1),
             label = avg_travel_time$describe,
             fillColor = ~pal(S0802_C01_090E),
             stroke = FALSE,
             radius = 50,
             fillOpacity = 1) %>%
  addLegend(pal = pal,
            values = ~S0802_C01_090E,
            bins = 10,
            opacity = 0.7, title = "Average Travel Time to Work (Minutes)",
            position = "topright")

tt_points_map
```


## Continuous Surface (Interpolation) Map

```{r interpolatoin map, message=FALSE, warning=FALSE}
travel_time_points_sp <- travel_time_points %>%
  st_transform(MA_state_plane) %>%
  as_Spatial()

tt_poly_sp <- avg_travel_time %>%
  st_transform(MA_state_plane) %>%
  as_Spatial()

boston_raster <- raster(tt_poly_sp, res=10)

gs <- gstat(formula = S0802_C01_090E~1, locations = travel_time_points_sp)
idw_interp <- interpolate(boston_raster, gs)

idw_interp_clip <- mask(idw_interp, tt_poly_sp)

tt_interpolation_map <- leaflet(travel_time_points) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addRasterImage(idw_interp_clip, colors = pal, opacity = 0.8) %>%
  addLegend(pal = pal,
            values = ~S0802_C01_090E,
            bins = 10,
            opacity = 0.7, title = "Average Travel Time to Work (Minutes)",
            position = "topright")

tt_interpolation_map
```













